### 2021-07-08 rad-icfp-2021/generalization.tex
.

\section{Beyond the Source language}
\label{sec:generalization}
In this section, we generalize our language by relaxing the imposed restrictions.
First, we allow free variables for the function argument of !reduce! in Section~\ref{sub:lift_reduce}. 
Then we show how to support more array operations (Section~\ref{sub:lift_more_arr_op}), conditionals (Section~\ref{sub:lift_conditional}), and non-smooth scalar operators (Section~\ref{sub:lift_non_smooth}).
Finally, we extend the language to support general arrays instead of supporting only $\Array{\reals}{n}$ in Section~\ref{sub:lift_gen_arr}.

@@

>>>>

\section{Beyond the Source language}
\label{sec:generalization}
In this section, we generalize our language by relaxing the imposed restrictions.
First, we allow free variables for the function argument of !reduce! in Section~\ref{sub:lift_reduce}. 
Then we show how to support more array operations (Section~\ref{sub:lift_more_arr_op}), conditionals (Section~\ref{sub:lift_conditional}), and non-smooth scalar operators (Section~\ref{sub:lift_non_smooth}).
Afterwards, we extend the language to support general arrays instead of supporting only $\Array{\reals}{n}$ in Section~\ref{sub:lift_gen_arr}. Finally, we provide a recipe for adding more constructs to the language in Section~\ref{sub:lift_recipe}.

@@

\subsection{Recipe for Adding More Constructs}
\label{sub:lift_recipe}
To add more constructs to our language, 
there is no need to go through UNF; most of the backend of our work can be used as a black box.
The non-smooth variables of the operator (typically booleans or integers) should be considered as external to the language, similarly to the way we treat the index $n$ of arrays.

\smartpara{Reverse AD}
If the operator has type $A\to B$, then one should provide its transpose Jacobian, a term $B\to A$. 
Such operators can often be unrolled to first-order programs.

\smartpara{Correctness}
To check correctness of the given Jacobian, it suffices to 
check that the semantics of the transpose Jacobian matches the Jacobian of the unrolled program.

\smartpara{Complexity}
To ensure the complexity guarantee of the whole transformation, one needs to check that there a linear usage of the continuation variable $Y$ and that the cost of the 
proposed Jacobian is at most $k$ times the complexity of the operator.
